%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Short Sectioned Assignment
% LaTeX Template
% Version 1.0 (5/5/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm} % Math packages

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
\fancyfoot[R]{\thepage} % Page numbering for right footer
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

%\setlength\parindent{5pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text

%----------------------------------------------------------------------------------------
%			ADDED PACKAGES, ETC.
%----------------------------------------------------------------------------------------
\usepackage{setspace} % Allows for double spacing, etc.
\usepackage{listings} % for code blocks
\usepackage{color}
 
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
 
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}
 
\lstset{style=mystyle}

\usepackage{hyperref} % links within document
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
\usepackage{array}
\newcommand*{\vertbar}{\rule[-1ex]{0.5pt}{2.5ex}}
\newcommand*{\horzbar}{\rule[.5ex]{2.5ex}{0.5pt}}

\usepackage{graphicx}% allows figures
\graphicspath{ {./images/} } 
\usepackage[margin=1in]{geometry}
\usepackage{bm} % allows bold
\usepackage{enumitem}

\DeclareMathOperator{\EX}{\mathbb{E}}% expected value
 
%----------------------------------------------------------------------------------------
%				GLOSSARY
%----------------------------------------------------------------------------------------
% Type these commands at the command line after editing glossary:
% pdflatex summary
% makeglossaries summary
% pdflatex summary

\usepackage{glossaries}
\makeglossaries
\loadglsentries{glossary}
\glsaddall % makes sure all definitions are in glossary even if they aren't mentioned in text

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{	
\normalfont \normalsize 
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Data Science Toolkit \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Matt Goodwin} % Your name

\date{\normalsize\today} % Today's date or a custom date
\onehalfspacing


%----------------------------------------------------------------------------------------
%	BEGIN DOCUMENT
%----------------------------------------------------------------------------------------

\begin{document}

% Print the title
\maketitle 

% Table of Contents
\tableofcontents
\newpage

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------

\section{Statistical Modeling Overview and Basic Theory}
\input{statistical_modeling_overview}

\section{Assessment of Models}
\input{metrics}

\section{Generalized Linear Models}
\input{glms}

\section{Bayesian Statistics}
\input{bayes_hierarchical_models}

\section{Ensemble}
\input{ensembles}

\section{Neural Networks}
\input{neural_nets}

\section{Unsupervised}
\input{pca}



\section{Basic Probability}
\input{probability}

\section{Classical Statistics}
\input{classical_statistics}

\section{Natural Language Processing}
\input{NLP}

\subsection{Problems}
\input{problems}

\section{Terms and Notation}

\subsection{Variable notiation}
\label{sec:notation}



Below explains notation used commonly when setting-up machine learning models and is taken from ESLII. Note that all vectors are assumed to be column vectors. To help understand the notation I use the example of predicting the sales of ice cream cones. 
\vspace{2mm}

\begin{itemize}
\item $X$ - represents an input variable. Even though input variable implies a single variable this could also be a vector. If we wanted to access a single variable from the input vector then we use notation $X_j$. So for example $X$ could include variables that describe the temperature ($X_{j}$), time or year ($X_{j+1}$), etc. 
\item $Y$ - represents a \emph{quantitative} output variable. This could be the sales of ice cream cones in dollars.
\item $G$ - represents a \emph{qualitative} output variable. This could be if we sale over 50 ice cream cones for example (yes or no). 
\item $x_i$ - represents an observed value of the variable $X$. Again this could be a vector. So to get the observed scalar value of the temperature for example we would write $x_{ij}$. 
\item $\bm{X}$ - matrix typically with dimensions $Nxp$. 
\item $\bm{x_j}$ -  in general vectors are not bold unless the distinction is being made that this is the vector of all observation on $X_j$. So $\bm{x_j}$ is of length $N$ and $x_i$ is of length $p$.
\end{itemize}


\section{Basic Statistical Concepts}

\subsection{Inference}

Inference is referring to using data to figure out the underlying properties of a population (which in turn allows us to understand the relationship between variables). I've been thinking of inference as referring to the process to understand the relationship between variables in a linear regression, but I think this is too narrow of a view. For example, if we look at using a t-test to compare two samples what we are really doing is using the data to estimate what the two distributions are that the data comes from and then determine if that is reasonable or not. TODO: How do the various techniques in statistics fit in with this idea of finding the parameters of the underlying data?


\clearpage
\printglossaries

\end{document}